---
output:
  html_document: default
  pdf_document: default
---
# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r} 
rdata <- read.csv("activity.csv")
rdata$date <- as.Date(rdata$date)

```


## What is mean total number of steps taken per day?
```{r}
gr <- split(rdata, rdata$date)
spd <- sapply(gr, function(x){ sum (x[,1], na.rm = TRUE)})
mnspd <- mean(spd)
mdspd <- median(spd)
hist(spd)
```


Mean steps taken per day: `r mnspd`. Median steps taken per day: `r mdspd`.


## What is the average daily activity pattern?
```{r}
gr <- split(rdata, rdata$interval)
apin <- sapply(gr, function(x){ mean (x[,1], na.rm = TRUE)})
plot(names(apin), apin, type = "l", xlab = "Interval", ylab = "Steps")
```

## Imputing missing values
Missing values in the dataset: `r sum(!complete.cases(rdata))`

Try to filling missing values with mean for that 5-minute interval.


```{r}
ndata <- rdata
ndata[!complete.cases(ndata), 1] <- apin[as.character(ndata[!complete.cases(ndata), 3])]
gr <- split(ndata, ndata$date)
spd <- sapply(gr, function(x){ sum (x[,1], na.rm = TRUE)})
nmnspd <- mean(spd)
nmdspd <- median(spd)
hist(spd)
```

Mean steps taken per day: `r nmnspd`. Median steps taken per day: `r nmdspd`.
After filling missing values the dataset seems more concentrated about mean.
Difference between means: `r nmnspd - mnspd`. Difference berween medians: `r nmdspd - mdspd`

## Are there differences in activity patterns between weekdays and weekends?


```{r}
ndata$week_part <- factor(
  ifelse(as.integer(format(ndata$date, format = '%u')) %in% c(1:5),
         'weekday', 'weekend'))

```

Panels in base plot system

```{r}
par(mfrow = c(2, 1), oma=c(1,1,0,0), mar=c(4,4,2,2))

stepsWeekend <- aggregate(steps ~ interval, 
                          data = subset(ndata,week_part == "weekend"), 
                          FUN = mean)
stepsWeekday <- aggregate(steps ~ interval, 
                          data = subset(ndata,week_part == "weekday"), 
                          FUN = mean)
## add the first
plot(stepsWeekend, type = "l", xlab="interval", ylab="steps", main="Weekend")
## add the second
plot(stepsWeekday, type = "l", xlab="interval", ylab="steps", main="weekday")
```

Yes, there seem to be differences! More steps in the weekend.
